---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, useFancyQuotes = FALSE)
```

MiRNAss is a machine learning method specifically designed for pre-miRNA
prediction. It takes advantage of unlabeled sequences to improve the prediction
rates even when there are just a few positive examples, and when the negative
examples are unreliable or are not good representatives of its class. 
Furthermore, the method can automatically search for negative examples if the
user is unable to provide them. MiRNAss can find a good boundary to divide the
pre-miRNAs from other groups of sequences; it automatically optimizes the
threshold that defines the classes boundaries, and thus, it is robust to high
class imbalance. Each step of the method is scalable and can handle large
volumes of data.

The last version of the package can be found at Bioconductor. Also, the
development version of the package can be found at:
<https://github.com/cyones/miRNAss>

# Contact
- Cristian Yones <cyones(at)sinc.unl.edu.ar>
- sinc(i):  http://fich.unl.edu.ar/sinc/

Usage
-----
After install the package, load miRNAss with the following command:

```{r}
library('miRNAss')
```

The following command is the simplest way to execute miRNAss:

```{r, eval=FALSE}
miRNAss(features, labels)
```

Where:

* features: is a data frame with the features extracted from hairpyn sequences,
one sequence per row and one numeric feature per column. The hairpin sequences
can be extracted from a raw genome using HExtractor
(<https://sourceforge.net/projects/sourcesinc/files/hextractor/>),
and the features can be calculated using miRNAfe
(<https://sourceforge.net/projects/sourcesinc/files/mirnafe/> or
<http://fich.unl.edu.ar/sinc/blog/web-demo/mirnafe-full/>). The whole pipeline
consist on extract the hairpins from a raw genome in fasta format. The output,
also in fasta format is the input to miRNAfe, which as result gives a comma
separate file that can be easily load into R.

* labels: is a numeric vector where the i-th element has a value of 1 if it is
a well-known pre-miRNA, a -1 if it is not a pre-miRNA, and zero if it is an
unknown sequence that has to be classified (predicted) by the method.

The data provided with the package can be used to test miRNAss. This small
dataset is composed of a small set of features extracted from 1000 hairpins
randomly extracted from C. elegans hairpins. To use miRNAss with this dataset,
first construct the label vector with the CLASS column
```{r}
y = as.numeric(celegans$CLASS)*2 - 1
```

Remove some labels to make a test
```{r}
y[sample(which(y > 0),200)] = 0
y[sample(which(y < 0),700)] = 0
```

Take all the features but remove the label column
```{r}
x = subset(celegans, select = -CLASS)
```

Call miRNAss with default parameters
```{r}
p = miRNAss(x,y)
```

Calculate some performance measures
```{r}
SE = mean(p[ celegans$CLASS & y == 0] > 0)
SP = mean(p[!celegans$CLASS & y == 0] < 0)
cat('Sensitivity: ', SE, '\nSpecificity: ', SP, '\n')
```

For more help about all the parameters and a full example execute:
```{r, eval=FALSE}
help(miRNAss)
```
# Datasets and test scripts
A set of experiments and comparisons with other methods can be done. The
scripts and the data of these experiments are contained in the file miRNAss-experiments.zip that can be found in:

<https://sourceforge.net/projects/sourcesinc/files/mirnass/>

To run these tests, after unzip the file, set this directory as the working
directory and simply run each script with the function 'source':

```{r, eval=FALSE}
setwd('experiment_scripts')
source('2_delta-mirBase.R')
```

This will generate one csv file for each test in the 'results' folder. It is
important to point that most of these experiments are computationally expensive
and could take quite a while (about 40 minutes for the experiment
2_delta-mirBase.R in an intel i7 PC). You can plot the results executing:

```{r, eval=FALSE}
source('plotResults.R')
```

The figures will be saved in the folder 'results'.
